The Role of Ice Compositions for Snowlines and the C/N/O Ratios in Active Disks

### 2016-07-18 tex/draft2/ms_revised.tex
upd

\title{The Role of Ice Compositions and Morphology for Snowlines and the C/N/O Ratios in Active Disks}

@@

\begin{abstract}
The elemental compositions of planets define their chemistry, and could potentially be used as beacons for their formation location if the elemental gas and grain ratios of planet birth environments, i.e. protoplanetary disks, are well understood. In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, their ice binding environment, and the presence of snowlines of major volatiles at different distances from the central star. We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice morphology on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk. The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions. Ice morphology and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model. Observations that anchor snowline locations at different stages of planet formation are therefore key to develop C/N/O ratios as a probe of planet formation zones.

@@

Both in Solar system comets and in protoplanetary disks, volatile carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{lodders03}, \citealt{mumma11}, \citealt{oberg11}, \citealt{boogert15}). However, some fraction of carbon may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust at some disk locations. In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}. Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}).  Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}. Because of the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk. Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin. In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).

The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain morphology. Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the morphology of the icy grain mantles. Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices. This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).

@@

In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice morphology, as well as nitrogen-to-oxygen (N/O) ratios. This paper is organized as follows. In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I. We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain morphology and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}. We address the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.95\textwidth]{CNO_and_snowlines_2.eps}
%\vspace{-0.5in}
\caption{The total carbon, nitrogen and oxygen abundance in solids as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel). Relevant volatile snowlines are marked by the vertical dashed lines. The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores. The total grain abundance increases with semimajor axis as more and more species freeze out.} 
\label{fig:CNOstatic}
\end{figure*}

@@

\subsection{Snowlines in a Static Disk: The Importance of Ice Morphology}
\label{sec:snowlines}

As we note in Section \ref{sec:intro}, the disk volatile composition and the ice morphology determine the location of important snowlines. In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e. H$_2$O, CO$_2$, CO, N$_2$, and to a lesser extent, CH$_4$ and NH$_3$. %To quantify the effect of the presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$, we use measured median CH$_4$ and NH$_3$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}. 
Our standard model is based on the median ice abundances observed toward Solar-type protostars \citep{oberg11a}, which are $n_{\rm CO_2}=0.29 \times n_{\rm H_2O}$, $n_{\rm CO}=0.38 \times n_{\rm H_2O}$, $n_{\rm CH_4}=0.0555 \times n_{\rm H_2O}$ (hereafter CH$_4$-mid) and $n_{\rm NH_3}=0.055 \times n_{\rm H_2O}$ (hereafter NH$_3$-mid). Here $n_{\rm H_2O} \approx 10^{-4} \times n_{\rm H}$ is the total water abundance \citep{vandishoeck06}, with $n_{\rm H}$ the hydrogen abundance in the disk midplane. For CO, we also take into account that the observed CO ice only traces some of the CO reservoir due to its high volatility, and similarly to \citet{oberg11} and Paper I we set the total CO abundance to $0.9 \times 10^{-4} n_{\rm H}$. Finally, we assume that all nitrogen not found in NH$_3$ is in N$_2$ and assume a Solar nitrogen abundance, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}. In effect, this model assumes no chemical evolution between the protostellar and disk midplane stages. This is reasonable for material that accretes onto the disk at large radii \citep{visser09}, but may overestimate the contribution of the original volatiles to the total volatile budget in the innermost disk.

We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}. The binding energies of H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, 767 K and 2965 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}, \citealt{martin14}). For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}. Figure \ref{fig:CNOstatic} shows the resulting snowline locations, assuming CO and N$_2$ pure ices (top panel), and CO and N$_2$ in water dominated ices (bottom panel). The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.  As expected, the total grain abundance increases with semimajor axis, as more and more species freeze out. Freeze-out at the CO$_2$ and CO snowlines pulls more heavy elements into the grains than in the case of the H$_2$O snowline.  The CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure. This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ.

@@

In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances toward low-mass protostars. In this section we only consider pure ices. %Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.eps}
%\vspace{-0.5in}
\caption{The C/O ratio in gas (solid lines) and dust (dashed lines) as a function of semimajor axis in a static disk, assuming no carbon is present in the form of CH$_4$ (top panel), the median observed CH$_4$ abundance is assumed (middle panel), and the maximum observed CH$_4$ abundance is assumed (bottom panel). The C/O estimates are performed assuming that the CO ices are in pure form. The vertical dotted lines mark the snowline locations of the main C and O carriers. The horizontal dotted lines represent the stellar C/O value. The presence of methane only modestly increases the C/O ratio in gas between the CO$_2$ and CH$_4$ snowlines.} 
\label{fig:COstatic}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{N_O_ratio.eps}
%\vspace{-0.5in}
\caption{The N/O ratio in gas (solid lines) and dust (dashed lines) as a function of semimajor axis in a static disk, assuming no nitrogen is present in the form of NH$_3$ (top panel), the median observed NH$_3$ abundance is assumed (middle panel), and the maximum observed NH$_3$ abundance is assumed (bottom panel). The N/O estimates are performed assuming that the CO and N$_2$ ices are in pure form. The vertical dotted lines mark the snowline locations of the main C,O and N carriers. The horizontal dotted lines represent the average N/O value in the disk. The gas-phase N/O ratio is enhanced by a factor of two between the H$_2$O and CO$_2$ snowlines compared to its average value, and by a factor of three between the CO$_2$ and CO snowlines. The arrows mark a highly elevated N/O ratio in gas between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region. The presence of NH$_3$ moderately decreases the N/O ratio in gas between the NH$_3$ and CO$_2$ snowlines.} 
\label{fig:Nstatic}
\end{figure}

@@

Here we use the model of Section \ref{sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk. Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above. We assume there is no carbon in the form of CH$_4$. The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.05$ cm do not form true snowlines. If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics. The full range of potential CO snowlines taking into account both ice morphology and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference. This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{C_O_water_ice.eps}
%\vspace{-0.5in}
\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as pure ice (top panel) or as water dominated ices (bottom panel). The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar. The C/O ratio in a static disk (black lines) is shown for comparison. The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$
snowlines in the viscous disk, as the relative fluxes of the desorbed icy
particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details). The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%. Taken together, disk dynamics and ice morphology move the CO snowline inward by a factor of $\sim$7.} 
\label{fig:CO_ratio}
\end{figure}

@@

Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for the same range of initial particle sizes as in Figure \ref{fig:CO_ratio}, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk. For simplicity, we assume that all nitrogen is the form of N$_2$. This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio. The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, water} \approx 11$ AU for N$_2$ in water dominated ices. Thus for each case (pure versus water dominated ices), the N$_2$ snowline moves inward by up to 50\% due to disk dynamics. By taking into account both ice morphology and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference. Similarly to the case for CO, the N$_2$ snowline may be close to 10 AU for certain disk models.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.eps}
%\vspace{-0.5in}
\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as pure ices (top panel) or as water dominated ices (bottom panel). The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar. The N/O ratio in a static disk (black lines) is shown for comparison. The arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details). Radial drift and gas accretion move the N$_2$ snowline inward by up to $\sim$50\% compared to a static disk. The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%. Taken together, disk dynamics and ice morphology move the N$_2$ snowline inward by a factor of $\sim$7. The results of an enhanced gas-phase N/O ratio between the H$_2$O and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}  
\label{fig:NO_ratio}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{CN_norm.eps}
%\vspace{-0.5in}
\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in a static disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$. The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines. The arrow indicates that the N/O ratio is enhanced by orders of magnitude compared to its average value between the CO and N$_2$ snowlines. The gaseous N/O ratio is enhanced throughout most of the disk, and more enhanced than the C/O ratio.}  
\label{fig:enhance}
\end{figure}

>>>>

\title{The Role of Ice Compositions for Snowlines and the C/N/O Ratios in Active Disks}

@@

\begin{abstract}
The elemental compositions of planets define their chemistry, and could potentially be used as beacons for their formation location if the elemental gas and grain ratios of planet birth environments, i.e. protoplanetary disks, are well understood. In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, their ice binding environment, and the presence of snowlines of major volatiles at different distances from the central star. We explore the effects of disk dynamical processes, molecular compositions and abundances, and ice \textbf{compositions} on the snowline locations of the main C, O and N carriers, and the C/N/O ratios in gas and dust throughout the disk. The gas-phase N/O ratio enhancement in the outer disk (exterior to the H$_2$O snowline) exceeds the C/O ratio enhancement for all reasonable volatile compositions. Ice \textbf{compositions} and disk dynamics individually change the snowline location of N$_2$, the main nitrogen carrier, by a factor of 2-3, and when considered together the range of possible N$_2$ snowline locations is $\sim$11-$\sim$79 AU in a standard disk model. Observations that anchor snowline locations at different stages of planet formation are therefore key to develop C/N/O ratios as a probe of planet formation zones.

@@

Both in Solar system comets and in protoplanetary disks, volatile carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{lodders03}, \citealt{mumma11}, \citealt{oberg11}, \citealt{boogert15}). However, some fraction of carbon may also be carried by CH$_4$ (e.g., \citealt{oberg08}), which may change the C/O ratio in gas and in dust at some disk locations. In the case of nitrogen, chemical models of the protostellar nebula (e.g., \citealt{owen01}) and of protoplanetary disks (e.g., \citealt{rodgers02}) suggest that N$_2$ was the dominant form of nitrogen, and that giant planets have accreted their nitrogen content primarily as N$_2$ \citep{mousis14}. Observations of Solar system bodies such as Titan and Pluto show that N$_2$ is prevalent in their atmospheres (\citealt{cruikshank93}, \citealt{owen93}), \textbf{although Titan's N$_2$ atmospheric composition is not primordial but rather originates from primordial NH$_3$ accreted by the satellite (\citealt{atreya78}, \citealt{mandt14}).}  %Moreover, the Rosetta spacecraft has recently made the first direct measurement of the N$_2$ abundance in comet 67P/Churyumov-Gerasimenko \citep{rubin15}. 
%\textbf{While CO and N$_2$ have similar condensation temperatures (e.g., \citealt{fayolle16}), their snowline locations may }  
Due to the high volatility of N$_2$, the gas phase nitrogen-to-oxygen (N/O) ratio in the outer disk may be even more enhanced than the C/O ratio compared to its average value in the disk. \textbf{This effect is most pronounced between the CO and N$_2$ snowlines, where the gas phase N/O ratio is enhanced by many orders of magnitude compared to the stellar value due to the depletion of oxygen gas in this region (see Section \ref{sec:results}).} Giant planets that form at wide separations should thus have an excess of nitrogen in their atmospheres, which could be used to trace their formation origin. In addition to N$_2$, a fraction of the nitrogen abundance may also be carried by less volatile species such as NH$_3$ (\citealt{bottinelli10}, \citealt{mumma11}).

The snowline locations of the main carbon, oxygen and nitrogen carriers strongly depend on the ice grain \textbf{composition}. Very volatile species, such as CO and N$_2$, present binding energies, and therefore snowline locations, that are sensitive to the details of the \textbf{composition} of the icy grain mantles. 
\textbf{Spectroscopic observations suggest that up to 90\% of the CO ice is frozen in a layer that is thick enough and separated from the H$_2$O ice layer underneath so that it can be considered pure ice (e.g., \citealt{pontoppidan03}). However, if the CO (or N$_2$) ice layer is thin enough, then it will interact with the H$_2$O ice substrate (e.g., \citealt{collings03}). The ice binding energy will be different in this water dominated environment than in the pure ice case.} Laboratory experiments (\citealt{collings03}, \citealt{oberg05}, \citealt{bisschop06}, \citealt{fayolle16}) have shown that CO and N$_2$ have significantly different binding energies depending on whether they are pure or water dominated ices. This implies that ices in different environments will sublimate at different radii, which will substantially change the disk regions where these volatiles are present in gaseous or solid form (see Section \ref{sec:static}).  \textbf{In protostellar cores, H$_2$O ice is primarily amorphous (e.g., \citealt{williams02}, \citealt{vandishoeck14}), and can either be compact or porous.  When temperatures exceed $\sim$80-90 K, H$_2$O ice acquires a crystalline structure \citep{schegerer10}. The CO binding energy is larger in an amorphous porous water ice environment than in the compact case \citep{fayolle16} and slightly lower than in the crystalline case \citep{noble12}. Here we consider pure ices and ices dominated by amorphous porous H$_2$O ice. We refer to the latter simply as water dominated ices unless noted otherwise.}

@@

In this work, we expand the coupled drift-desorption model developed in (\citealt{piso15b}; hereafter Paper I) by considering additional volatile molecules and abundances, ice \textbf{compositions}, as well as nitrogen-to-oxygen (N/O) ratios. This paper is organized as follows. In Section \ref{sec:review}, we review the drift-desorption model developed in Paper I. We discuss the effect of different abundances of the main carbon, oxygen and nitrogen carriers, grain \textbf{compositions} and disk dynamics on snowline locations and the C/N/O ratios in Section \ref{sec:results}. We address the implications of our results in Section \ref{sec:discussion} and summarize our findings in Section \ref{sec:summary}.

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.95\textwidth]{CNO_and_snowlines_2.pdf}
%\vspace{-0.5in}
\caption{The total carbon, nitrogen and oxygen abundance in solids as a function of semimajor axis in a static disk, for CO and N$_2$ as pure ices (top panel) and water dominated ices (bottom panel). Relevant volatile snowlines are marked by the vertical dashed lines. The grain abundances are calculated as a function of the observed median CH$_4$ and NH$_3$ abundances in protostellar cores. The total grain abundance increases with semimajor axis as more and more species freeze out.} 
\label{fig:CNOstatic}
\end{figure*}

@@

\subsection{Snowlines in a Static Disk: The Importance of Ice \textbf{Compositions}}
\label{sec:snowlines}

As we note in Section \ref{sec:intro}, the disk volatile composition and the ice \textbf{composition} determine the location of important snowlines. In this work we focus on the primary carbon, oxygen and nitrogen carriers, i.e. H$_2$O, CO$_2$, CO, N$_2$, and to a lesser extent, CH$_4$ and NH$_3$. %To quantify the effect of the presence of some carbon in the form of CH$_4$ and of some nitrogen in the form of NH$_3$, we use measured median CH$_4$ and NH$_3$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}. 
Our standard model is based on the median ice abundances observed toward Solar-type protostars \citep{oberg11a}, which are $n_{\rm CO_2}=0.29 \times n_{\rm H_2O}$, $n_{\rm CO}=0.38 \times n_{\rm H_2O}$, $n_{\rm CH_4}=0.0555 \times n_{\rm H_2O}$ (hereafter CH$_4$-mid) and $n_{\rm NH_3}=0.055 \times n_{\rm H_2O}$ (hereafter NH$_3$-mid). Here $n_{\rm H_2O} \approx 10^{-4} \times n_{\rm H}$ is the total water abundance \citep{vandishoeck06}, with $n_{\rm H}$ the hydrogen abundance in the disk midplane. For CO, we also take into account that the observed CO ice only traces some of the CO reservoir due to its high volatility, and similarly to \citet{oberg11} and Paper I we set the total CO abundance to $0.9 \times 10^{-4} n_{\rm H}$. Finally, we assume that all nitrogen not found in NH$_3$ is in N$_2$ and assume a Solar nitrogen abundance, $n_{\rm N}=8 \times 10^{-5} n_{\rm H}$ \citep{lodders03}. In effect, this model assumes no chemical evolution between the protostellar and disk midplane stages. This is reasonable for material that accretes onto the disk at large radii \citep{visser09}, but may overestimate the contribution of the original volatiles to the total volatile budget in the innermost disk.

We determine the location of the H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}. The binding energies of H$_2$O, CO$_2$, CO, CH$_4$, N$_2$ and NH$_3$ as pure ices are 5800 K, 2000 K, 834 K, 1300 K, 767 K and 2965 K, respectively (\citealt{fraser01}, \citealt{collings04}, \citealt{fayolle16}, \citealt{garrod06}, \citealt{martin14}). \textbf{For our fiducial disk model, these energies correspond to disk temperatures of 143 K, 48 K, 21 K, 30 K, 18 K and 68 K, respectively.} For CO and N$_2$ as water dominated ices, the binding energies are 1388 K and 1266 K, respectively \citep{fayolle16}, \textbf{corresponding to disk temperatures of 34 K and 31 K, respectively}. Figure \ref{fig:CNOstatic} shows the resulting snowline locations, assuming CO and N$_2$ pure ices (top panel), and CO and N$_2$ in water dominated ices (bottom panel). The ordinate displays the total carbon, oxygen and nitrogen abundance in solids as a function of the hydrogen total abundance.  As expected, the total grain abundance increases with semimajor axis, as more and more species freeze out. Freeze-out at the CO$_2$ and CO snowlines pulls more heavy elements into the grains than in the case of the H$_2$O snowline.  The CO and N$_2$ snowlines move several tens of AU inward if the ices are water dominated rather than pure. This changes the chemical abundances both in gas and dust throughout the disk, directly affecting the compositions of nascent giant planets forming in situ. %\textbf{We note that the CO and N$_2$ snowlines will move inward if the ices are in an amorphous compact water environment and slightly shift outward in the case of a crystalline H$_2$O ice substrate.}

@@

In this section we determine the C/O and N/O ratios in gas and dust throughout our static disk, and to what extent they are affected by the presence of CH$_4$ and NH$_3$ over the full range of observed CH$_4$ and NH$_3$ abundances toward low-mass protostars. In this section we only consider pure ices, \textbf{i.e. ices that are layered on a silicate mantle. In reality, the icy grain will have a layered structure with volatiles residing on top of a H$_2$O ice substrate. However, if the volatile ice layer is thick enough and separated from the H$_2$O ice layer, the interaction between H$_2$O and the other volatile species will be minimal and thus the ices can be considered pure rather than water dominated (see also Section \ref{sec:intro}).}

 %Since here we are primarily interested in the magnitude of the C/N/O ratios rather than snowline locations (which have been discussed in Section \ref{sec:snowlines}), we assume for simplicity that CO and N$_2$ are pure ices.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{C_O_ratio_CH4.pdf}
%\vspace{-0.5in}
\caption{The C/O ratio in gas (solid lines) and dust (dashed lines) as a function of semimajor axis in a static disk, assuming no carbon is present in the form of CH$_4$ (top panel), the median observed CH$_4$ abundance is assumed (middle panel), and the maximum observed CH$_4$ abundance is assumed (bottom panel). The C/O estimates are performed assuming that the CO ices are in pure form. The vertical dotted lines mark the snowline locations of the main C and O carriers. The horizontal dotted lines represent the stellar C/O value. The presence of methane only modestly increases the C/O ratio in gas between the CO$_2$ and CH$_4$ snowlines.} 
\label{fig:COstatic}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{N_O_ratio.pdf}
%\vspace{-0.5in}
\caption{The N/O ratio in gas (solid lines) and dust (dashed lines) as a function of semimajor axis in a static disk, assuming no nitrogen is present in the form of NH$_3$ (top panel), the median observed NH$_3$ abundance is assumed (middle panel), and the maximum observed NH$_3$ abundance is assumed (bottom panel). The N/O estimates are performed assuming that the CO and N$_2$ ices are in pure form. The vertical dotted lines mark the snowline locations of the main C,O and N carriers. The horizontal dotted lines represent the average N/O value in the disk. The gas-phase N/O ratio is enhanced by a factor of two between the H$_2$O and CO$_2$ snowlines compared to its average value, and by a factor of three between the CO$_2$ and CO snowlines. The arrows mark a highly elevated N/O ratio in gas between the CO and N$_2$ snowlines due to the depletion of oxygen gas in this region. The presence of NH$_3$ moderately decreases the N/O ratio in gas between the NH$_3$ and CO$_2$ snowlines.} 
\label{fig:Nstatic}
\end{figure}

@@

Here we use the model of Section \ref{sec:review} to estimate the movement of the CO and N$_2$ snowlines for different grain morphologies in a viscous disk. Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above. We assume there is no carbon in the form of CH$_4$. The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.05$ cm do not form true snowlines. If the CO binding environment is known, the CO snowline moves inward by up to $\sim$50 \% compared to a static disk for each case (pure and water dominated ices) due to disk dynamics. The full range of potential CO snowlines taking into account both ice \textbf{compositions} and disk dynamics span $\sim$8.7 AU to $\sim$61 AU, which is a factor of $\sim$7 difference. This implies that gas phase C/O ratios of order unity may be reached in the giant planet forming zone, and the CO snowline may be inside 10 AU for certain disk parameters.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{C_O_water_ice.pdf}
%\vspace{-0.5in}
\caption{C/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO as pure ice (top panel) or as water dominated ices (bottom panel). The H$_2$O, CO$_2$ and CO snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar. The C/O ratio in a static disk (black lines) is shown for comparison. The arrows show that the C/O ratio in gas will decrease inside the H$_2$O and CO$_2$
snowlines in the viscous disk, as the relative fluxes of the desorbed icy
particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details). The presence of CO in a water ice environment rather than as pure ice moves the CO snowline significantly inward by $\sim$70\%. Taken together, disk dynamics and ice \textbf{compositions} move the CO snowline inward by a factor of $\sim$7.} 
\label{fig:CO_ratio}
\end{figure}

@@

Figure \ref{fig:NO_ratio} shows the H$_2$O, CO$_2$, CO and N$_2$ snowline locations in a viscous disk for the same range of initial particle sizes as in Figure \ref{fig:CO_ratio}, and with the CO and N$_2$ snowlines calculated assuming different grain morphologies as explained above, as well as estimates for the N/O ratio throughout the disk. For simplicity, we assume that all nitrogen is the form of N$_2$. This choice is justified since the presence of some NH$_3$ only moderately changes the N/O ratio (see Figure \ref{fig:Nstatic}), and since we are primarily interested in the N$_2$ snowline locations rather than exact values for the N/O ratio. The innermost N$_2$ snowlines in the viscous disk, created by particles with $s \sim 7$ m for our fiducial model, are located at $r_{\rm N_2, pure} \approx 42$ AU for N$_2$ as pure ice and at $r_{\rm N_2, water} \approx 11$ AU for N$_2$ in water dominated ices. Thus for each case (pure versus water dominated ices), the N$_2$ snowline moves inward by up to 50\% due to disk dynamics. By taking into account both ice \textbf{compositions} and disk dynamics, the full range of potential N$_2$ snowlines span $\sim$11 to $\sim$79 AU, which is a factor of $\sim$7 difference. Similarly to the case for CO, the N$_2$ snowline may be close to 10 AU for certain disk models.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{N_O_water_ice_many.pdf}
%\vspace{-0.5in}
\caption{N/O ratio estimates in gas (solid lines) and dust (dashed lines) as function of semimajor axis in a viscous disk, for CO and N$_2$ as pure ices (top panel) or as water dominated ices (bottom panel). The H$_2$O, CO$_2$, CO and N$_2$ snowlines are shown for particles with initial sizes $\sim0.05$ cm $\lesssim s \lesssim$ 7 m as indicated by the color bar. The N/O ratio in a static disk (black lines) is shown for comparison. The arrows show that the N/O ratio in gas will decrease inside the H$_2$O and CO$_2$ snowlines in the viscous disk, as the relative fluxes of the desorbed icy
particles and the overall nebular gas will cause an excess of oxygen gas inside these snowlines (see Paper I for details). Radial drift and gas accretion move the N$_2$ snowline inward by up to $\sim$50\% compared to a static disk. The presence of N$_2$ in a water ice environment rather than as pure ice moves the N$_2$ snowline significantly inward by $\sim$70\%. Taken together, disk dynamics and ice \textbf{compositions} move the N$_2$ snowline inward by a factor of $\sim$7. The results of an enhanced gas-phase N/O ratio between the H$_2$O and CO snowlines compared to its average value, and of highly elevated N/O ratios in gas between the CO and N$_2$ snowlines (see Figure \ref{fig:Nstatic}), are preserved.}  
\label{fig:NO_ratio}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{CN_norm.pdf}
%\vspace{-0.5in}
\caption{Gas phase C/O (blue curve) and N/O (red curve) ratios divided by the average C/O and N/O ratio in a static disk, assuming CO and N$_2$ are pure ices, and there is no CH$_4$ or NH$_3$. The dashed vertical lines mark the H$_2$O, CO$_2$, CO and N$_2$ snowlines. The arrow indicates that the N/O ratio is enhanced by orders of magnitude compared to its average value between the CO and N$_2$ snowlines. The gaseous N/O ratio is enhanced throughout most of the disk, and more enhanced than the C/O ratio.}  
\label{fig:enhance}
\end{figure}

Due to disk dynamics and ice \textbf{compositions}, the locations of the CO and N$_2$ snowlines, and thus the disk regions with highly elevated gas phase N/O and C/O ratios, are uncertain and may span tens of AU. Both ice morphologies discussed in this study, pure and water dominated ices, are plausible in protoplanetary disks and depend on whether H$_2$O and CO ices formed on similar timescales or successively (e.g., \citealt{garrod11}). Observations of protostellar cores show that a large fraction of CO is bound in a pure ice multilayer \citep{pontoppidan03}, but theoretical models also suggest an icy mantle structure where CO resides on a H$_2$O ice layer (e.g., \citealt{collings03}). One can also imagine a scenario where CO is in a water binding environment and N$_2$ is not. This could be attributed to the fact that H$_2$O may bind preferentially to CO than N$_2$, since both H$_2$O and CO are polar molecules while N$_2$ is not. It is also possible for N$_2$ ices to form later than CO (e.g., \citealt{pagani12}), and thus be deposited on the outer layers of the icy mantles which are typically water poor (e.g., \citealt{garrod11}). The impact of the ice environment on the snowline location is much smaller in the case of CO$_2$ and NH$_3$, as their binding energies and behavior are closer to that of H$_2$O. No detailed measurements for the CH$_4$ binding energy in a water environment exist so far, but due to its low desorption temperature a similar behavior to that of CO and N$_2$ would be expected. While the presence of some carbon in the form of CH$_4$ only modestly affects our results, CH$_4$ may become important in disks where a large fraction of the CO abundance has been converted into hydrocarbons (e.g., \citealt{du15}).
